---
title: "Crane_trachea (Body mass conceptual figure)"
author: "Chauncey Gadek and Trinity Casaus" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 5
    code_folding: hide
    #df_print: paged
    #df_print: kable
    #toc_float: true
      #collapsed: false
      #smooth_scroll: TRUE
    theme: cosmo #spacelab #yeti #united #cosmo
    highlight: tango
  pdf_document:
    df_print: kable
fontsize: 12pt
geometry: margin=0.25in
always_allow_html: yes
bibliography: references.bib  
---

```{=html}
<style>
/* HTML FORMATTING */
h1, .h1, h2, .h2, h3, .h3, h4, .h4, h5, .h5 {
  margin-top: 25px; /* space before each header */
  font-weight: bold; /* bold headers */
}
</style>
```
```{R, echo=FALSE}
# I set some GLOBAL R chunk options here.
#   (to hide this message add "echo=FALSE" to the code chunk options)
rm(list =ls (all = TRUE)) #This removes objects from global environ
knitr::opts_chunk$set(echo=T, comment = NA, message = FALSE, warning = FALSE, width = 100)
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 6)
#setwd("C:/Users/Trinity's/Dropbox/Crane_trachea")
#setwd("~/Dropbox/Crane_trachea") #need to change to local directory
```

# Load Packages

```{R}
library(reshape)
library(reshape2)
library(plyr)
library(dplyr)
library(tidyr)
library(GGally)
library(gridExtra)
library(gplots)
library(ggplot2)
library(ggExtra)
library(cowplot)
library(viridis)
```

This script is to create a figure illustrating the variation of SSD in birds and a larger group of animals. I have downloaded data from @herberstein2022animaltraits who recently published an animal traits database including male and female mass for a handful of species. @Amniote2015 is an older dataset with ~ 5000 birds that have mean values of mass by sex. Their [metadata](https://www.esapubs.org/archive/ecol/E096/269/metadata.php) does not describe how they obtained the mass metrics, but the [abstract](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1890/15-0846R.1) suggest that they are median values. @Amniote2015 is a large dataset with containing reptiles birds and mammals. @lislevand2007avian is an older dataset with ~ 5000 birds that have mean values of mass by sex. Their [metadata](https://www.esapubs.org/archive/ecol/E096/269/metadata.php), does not describe how they obtained the mass metrics, but gives sample sizes at points so probably either mean or median. 

There is considerable redundancy in these datasets. After loading them and wrangling them into the same units and format, I will combine them and take the mean of all values by taxon. This will ensure that different measurements for the same species are further combined.

# Read in all kinds of data
```{r}
Amniote_2015 <- read.csv("data/Amniote_Database_Aug_2015.csv")
animal_traits <- read.csv("data/animaltraits.csv")
avian_Lislevand_2007 <- read.delim("data/avian_body_size.txt", sep="\t")

```

# Wrangle the diverse data
```{r}
Amniote_2015_ALL <- Amniote_2015%>%
  filter(male_body_mass_g >= -500,
         female_body_mass_g >= -500)%>%
  na.omit()%>%
  mutate(taxon = factor(paste(genus, species, sep=" ")),
         male = male_body_mass_g,
         female = female_body_mass_g,
         sex_diff_male_female = male - female,
         SDI = (male-female)/female)%>%
  dplyr::select(taxon, male, female, sex_diff_male_female, SDI)
Amniote_2015_ALL$taxon <- reorder(Amniote_2015_ALL$taxon, Amniote_2015_ALL$sex_diff_male_female)


animal_traits_all <- animal_traits%>%
  dplyr::select(genus, species, sex, body.mass)%>%
  mutate(body.mass = body.mass *1000, #change from kg to g
         taxon = factor(species))%>%
  filter(!is.na(body.mass) & !is.na(sex) & !sex == "")%>%
  group_by(taxon, sex)%>%
  summarise(body.mass = mean(body.mass, na.rm = T))%>%
  pivot_wider(.,id_cols =c(taxon), values_from = body.mass, names_from = sex)%>%
  dplyr::select(!both)%>%
  na.omit()%>%
  mutate(sex_diff_male_female = male - female,
         SDI = (male-female)/female)%>%
  dplyr::select(taxon, male, female, sex_diff_male_female, SDI)

Amniote_2015_birds <- Amniote_2015%>%
  filter(class == "Aves",
         male_body_mass_g >= -500,
         female_body_mass_g >= -500)%>%
  na.omit()%>%
  mutate(male = male_body_mass_g,
         female = female_body_mass_g,
          taxon = factor(paste(genus, species, sep=" ")),
         sex_diff_male_female = male - female,
         SDI = (male-female)/female)%>%
  dplyr::select(taxon, male, female, sex_diff_male_female, SDI)


animal_traits_birds <- animal_traits%>%
  filter(class=="Aves")%>%
  dplyr::select(genus, species, sex, body.mass)%>%
  mutate(body.mass = body.mass *1000, #change from kg to g
         taxon = factor(species))%>%
  filter(!is.na(body.mass) & !is.na(sex) & !sex == "")%>%
  group_by(taxon, sex)%>%
  summarise(body.mass = mean(body.mass, na.rm = T))%>%
  pivot_wider(.,id_cols =c(taxon), values_from = body.mass, names_from = sex)%>%
  na.omit()%>%
  mutate(sex_diff_male_female = male - female,
         SDI = (male-female)/female)%>%
  dplyr::select(taxon, male, female, sex_diff_male_female, SDI)

Livesand_birds <- avian_Lislevand_2007%>%
  dplyr::select(Species_name, M_mass, F_mass)%>%
  mutate(taxon = factor(Species_name))%>%
  filter(M_mass >= -500,
         F_mass >= -500)%>%
  na.omit()%>%
  group_by(taxon)%>%
  summarise(male = mean(M_mass, na.rm = T),
            female = mean(F_mass, na.rm = T))%>%
  mutate(sex_diff_male_female = male - female,
         SDI = (male-female)/female)%>%
  dplyr::select(taxon, male, female, sex_diff_male_female, SDI)

#combine datasets
all_animals <- rbind(Amniote_2015_ALL, animal_traits_all)
all_birds <- rbind(Amniote_2015_birds, animal_traits_birds, Livesand_birds)

#remove duplicates
all_animals <- all_animals%>%
  distinct(., taxon, .keep_all = TRUE)%>%
  droplevels()
all_birds <- all_birds%>%
  distinct(., taxon, .keep_all = TRUE)%>%
  droplevels()

#add crane no crane column to birds
all_birds <- all_birds%>%
  mutate(crane_y_n = factor(if_else(taxon %in% c("Grus canadensis canadensis", "Grus canadensis tabida"), 1,0)))

#Reorder by sex diff for plotting
all_animals$taxon <- reorder(all_animals$taxon, all_animals$sex_diff_male_female)
all_birds$taxon <- reorder(all_birds$taxon, all_birds$sex_diff_male_female)
```

# Plot
## All Amniotes
```{r}
all_animals %>%
  mutate(sex.diff.log = log10(male)-log10(female),
         sex.diff = log10(male)-log10(female))%>%
    arrange(taxon, sex_diff_male_female) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon)))%>% # reset your factor-column based on that order
ggplot(., aes(x= taxon, y=log10(sex_diff_male_female), color=log10(sex_diff_male_female), fill=log10(sex_diff_male_female)))+
  geom_col(width=0.1)+
  geom_col(
     data = all_birds %>%
  #filter(sex_diff_male_female > -0.7) %>%
  #mutate(sex.diff.log = log10(male)-log10(female))%>%
  filter(crane_y_n == 1)%>%
    arrange(taxon, sex_diff_male_female),
  fill = "black",
  color = "black",
  width = 0.01
  ) +
  scale_colour_viridis_c()+
   # scale_fill_gradient2(low = "#d1495b", mid = "#fde09f", high = "#00798c", midpoint = 0, breaks = c(-0.0001,0,0.001))+
   # scale_colour_gradient2(low = "#d1495b", mid = "#fde09f", high = "#00798c", midpoint = 0, breaks = c(-0.0001,0,0.001))+
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

```

```{r}
Amniote_2015_ALL %>%
  filter(SDI < 6) %>%
  mutate(sex.diff.log = log10(male) - log10(female)) %>%
  arrange(SDI) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon))) %>% # reset your factor-column based on that order
  ggplot(.
         ,
         aes(
           x = log10(male),
           y = SDI,
           fill = SDI,
           color = SDI
         )) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm",
              color = "grey57",
              size = 1) +
  geom_point(
    data = all_birds %>%
      #filter(sex_diff_male_female > -0.7) %>%
      mutate(sex.diff.log = log10(male) - log10(female)) %>%
      filter(crane_y_n == 1) %>%
      arrange(taxon, SDI),
    aes(),
    fill = "darkorchid",
    color = "darkorchid",
    alpha = 0.7,
    size = 3,
    shape=21
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_gradient2(
    low = "#d1492d",
    mid = "#fde09f",
    high = "#00798c"
  ) +
  scale_colour_gradient2(
    low = "#d1492d",
    mid = "#FFE6BE",
    high = "#00798c",
    midpoint = 0.1,
    breaks = c(-0.25, 0, 0.25, 0.5, 1, 1.5, 2)
  ) +
  theme(legend.title = element_blank(),
        legend.position = c(2, 9)) +
  theme_cowplot()
```

```{r}
Amniote_2015_ALL %>%
  filter(SDI < 6) %>%
  mutate(sex.diff.log = log10(male) - log10(female)) %>%
  arrange(SDI) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon))) %>% # reset your factor-column based on that order
  ggplot(.
         ,
         aes(
           x = log10(male),
           y = SDI,
           fill = SDI,
           color = SDI
         )) +
  geom_col(alpha = 0.5, binwidth=1) +
  geom_point(
    data = all_birds %>%
      #filter(sex_diff_male_female > -0.7) %>%
      mutate(sex.diff.log = log10(male) - log10(female)) %>%
      filter(crane_y_n == 1) %>%
      arrange(taxon, SDI),
    aes(),
    fill = "darkorchid",
    color = "darkorchid",
    alpha = 0.7,
    size = 3,
    shape=21
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_gradient2(
    low = "#d1492d",
    mid = "#fde09f",
    high = "#00798c"
  ) +
  scale_colour_gradient2(
    low = "#d1492d",
    mid = "#FFE6BE",
    high = "#00798c",
    midpoint = 0.1,
    breaks = c(-0.25, 0, 0.25, 0.5, 1, 1.5, 2)
  ) +
  theme(legend.title = element_blank(),
        legend.position = c(2, 9)) +
  theme_cowplot()
```

## Birds only
```{r}
all_birds %>%
  filter(sex_diff_male_female > -0.7) %>%
  mutate(sex.diff.log = log10(male)-log10(female))%>%
    arrange(sex.diff.log) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon)))%>% # reset your factor-column based on that order
  ggplot(.
         ,
         aes(x = taxon,
             y = sex.diff.log,
             fill = sex.diff.log,
             color = sex.diff.log)) +
  geom_col() +
  geom_col(
     data = all_birds %>%
  filter(sex_diff_male_female > -0.7) %>%
  mutate(sex.diff.log = log10(male)-log10(female))%>%
  filter(crane_y_n == 1)%>%
    arrange(taxon, sex.diff.log),
  fill = "black",
  color = "black",
  width = 0.01
  ) +
   scale_fill_gradient2(low = "#d1495b", mid = "#fde09f", high = "#00798c", midpoint = 0, breaks = c(-0.0001,0,0.001))+
   scale_colour_gradient2(low = "#d1495b", mid = "#fde09f", high = "#00798c", midpoint = 0, breaks = c(-0.0001,0,0.001))+
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
  

```

## Birds only SDI
```{r}


all_birds %>%
  arrange(SDI) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon)))%>% # reset your factor-column based on that order
  ggplot(.
         ,
         aes(x = taxon,
             y = (SDI*100),
             color = (SDI*100))) +
  geom_col(position = "dodge2") +
  geom_col(
    data = all_birds %>%
      filter(crane_y_n == 1),
    fill = "black",
    color = "black",
    width = 0.01
  ) +
#   scale_color_gradient2(
#   low = munsell::mnsl("5B 7/8"),
#   high = munsell::mnsl("5Y 7/8"),
#   mid = munsell::mnsl("N 7/0"),
#   midpoint = .02
# ) +
  scale_fill_gradient2(low = "#d1495b", mid = "#fde09f", high = "#00798c", midpoint = 0, breaks = c(-2,0,2))+
   scale_colour_gradient2(low = "#d1495b", mid = "#fde09f", high = "#00798c", midpoint = 0, breaks = c(-0.2,0,0.2))+
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
  

```


## Birds only
```{r}
all_birds %>%
  filter(sex_diff_male_female > -0.7) %>%
  mutate(sex.diff.log = log10(male)-log10(female))%>%
    arrange(sex.diff.log) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon)))%>% # reset your factor-column based on that order
  ggplot(.
         ,
         aes(x = log10(male),
             y = sex.diff.log,
             fill = SDI,
             color = SDI)) +
  geom_point(alpha=0.5) +
  geom_point(
     data = all_birds %>%
  filter(sex_diff_male_female > -0.7) %>%
  mutate(sex.diff.log = log10(male)-log10(female))%>%
  filter(crane_y_n == 1)%>%
    arrange(taxon, sex.diff.log),
  aes(),
  fill = "darkorchid",
  color = "darkorchid",
  alpha=0.7,
  size=3
  
  ) +
  geom_hline(yintercept=0, linetype="dashed")+
   scale_fill_gradient2(low = "#d1492d", mid = "#fde09f", high = "#00798c", midpoint = 0.4, breaks=c(0.3, 0.5))+
   scale_colour_gradient2(low = "#d1492d", mid = "#fde09f", high = "#00798c", midpoint = 0.4,breaks=c(0.3, 0.5))+
  theme(
    legend.title = element_blank()
  )+
  theme_cowplot()
  

```

```{r}
all_birds %>%
  #filter(sex_diff_male_female > -0.7) %>%
  mutate(sex.diff.log = log10(male) - log10(female)) %>%
  arrange(SDI) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon))) %>% # reset your factor-column based on that order
  ggplot(.
         ,
         aes(
           x = log10(male),
           y = SDI,
           fill = SDI,
           color = SDI
         )) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm",
              color = "grey57",
              size = 1) +
  geom_point(
    data = all_birds %>%
      #filter(sex_diff_male_female > -0.7) %>%
      mutate(sex.diff.log = log10(male) - log10(female)) %>%
      filter(crane_y_n == 1) %>%
      arrange(taxon, SDI),
    aes(),
    fill = "darkorchid",
    color = "darkorchid",
    alpha = 0.7,
    size = 3,
    shape=21
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_gradient2(
    low = "#d1492d",
    mid = "#fde09f",
    high = "#00798c"
  ) +
  scale_colour_gradient2(
    low = "#d1492d",
    mid = "#FFE6BE",
    high = "#00798c",
    midpoint = 0.1,
    breaks = c(-0.25, 0, 0.25, 0.5, 1, 1.5, 2)
  ) +
  theme(legend.title = element_blank(),
        legend.position = c(2, 9)) +
  theme_cowplot()




```

# Density all birds
```{r}
all_birds %>%
  #filter(sex_diff_male_female > -0.7) %>%
  mutate(sex.diff.log = log10(male) - log10(female)) %>%
  arrange(SDI) %>%               # sort your dataframe
  mutate(taxon = factor(taxon, unique(taxon))) %>% # reset your factor-column based on that order
  ggplot(.
         ,
         aes(
           x = sex.diff.log,
           fill = SDI,
           color = SDI
         )) +
  geom_density(alpha = 0.5) +
  geom_point(
    data = all_birds %>%
      #filter(sex_diff_male_female > -0.7) %>%
      mutate(sex.diff.log = log10(male) - log10(female)) %>%
      filter(crane_y_n == 1) %>%
      arrange(taxon, SDI),
    aes(y=0),
    fill = "darkorchid",
    color = "darkorchid",
    alpha = 0.7,
    size = 3,
    shape=21
  ) +
  scale_fill_gradient2(
    low = "#d1492d",
    mid = "#fde09f",
    high = "#00798c"
  ) +
  theme(legend.title = element_blank(),
        legend.position = c(2, 9)) +
  theme_cowplot()

```
# References